---
title: "Comparing GUESS with BaseQTL Jul23 - looking for beta cutoffs using Zscore"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "~/HRJ_monocytes/leo_triplets")
knitr::opts_knit$set(root.dir = "~/Documents/analysis/from_leo/comparisons")
library(dplyr)
library(tidyr)
library(stringr)
library(data.table)
library(ggplot2)
library(forcats)
library(confintr)
```

### Compare the effects from BaseQTL with the effects from GUESS
For each modality, first look at the normal approximation in BaseQTL versus the post.proc poster probability for the best model containing that SNP in GUESS.
Then look at the distributions of effects across both analyses.


```{r load datasets}
#trip <- fread("./results/dec22/filtered/PP_stats_triplet_SMALL_Dec22_DpnFilt.txt")
#peak_locs <- as.data.table(read.table("./input/alex_inputs/ATAC_gene_locations.csv"))
#names(peak_locs) = c("Peak_name", "chrom", "peakStart", "peakEnd")
#base_chic <- fread("~/HRJ_monocytes/BaseQTL/elena_results/CHi-C/round2_refBias/CHiC_BaseQTL_Elena_round2_wNormApprox.txt")
#dpn <- fread("~/HRJ_monocytes/findmotifs/find_new_dpnII/SNPs_and_LD_proxies_Jan2020.DpnII.marked.REFALTseqs.txt")
#base_atac <- fread("~/HRJ_monocytes/BaseQTL/elena_results/ATAC/round2_refBias/atac_stan.summary.txt")
#base_rna <- fread("~/HRJ_monocytes/BaseQTL/elena_results/RNA/p1.eqtl.summary.txt")

trip <- fread("../jul23/PP_stats_triplet_mPPI_FDR_0.05_SMALL/PP_stats_triplet_FDR_0.05_SMALL.txt")
peak_locs <- as.data.table(read.table("../input/alex_inputs/ATAC_gene_locations.csv"))
names(peak_locs) = c("Peak_name", "chrom", "peakStart", "peakEnd")
base_chic <- fread("../../BaseQTL_results/CHiC/round2_refBias/CHiC_BaseQTL_Elena_round2_wNormApprox.txt")
dpn <- fread("../../BaseQTL_results/CHiC/round2_refBias/SNPs_and_LD_proxies_Jan2020.DpnII.marked.REFALTseqs.txt")
base_atac <- fread("../../BaseQTL_results/ATAC/round2_refBias/atac_stan.summary.txt")
base_rna <- fread("../../BaseQTL_results/RNA/p0.99.eqtl.summary.txt")

### UPDATE 18th May 2023: take only distal sets {SEE END OF SCRIPT FOR CODE TO MAKE THIS}
distal_sets <- fread("../input/Helen_trulyDistalSets.txt")
distal_sets_small <- unique(distal_sets[, .(set, ENSG_Gene, ENSG_repGene)])

trip[, Gene := tstrsplit(`Triplet name`, split = "/", keep = 3)]

trip_distal1 <- trip[distal_sets_small, on = c(`Triplet ID` = "set", Gene = "ENSG_Gene"), nomatch = NULL]
trip_distal1[, ENSG_repGene := NULL]
trip_distal2 <- trip[distal_sets_small, on = c(`Triplet ID` = "set", Gene = "ENSG_repGene"), nomatch = NULL]
trip_distal2[, ENSG_Gene := NULL]

trip_distal <- unique(rbind(trip_distal1, trip_distal2))


# put tables next to each other
# have numbers next to graphs -no in different colours and the no in different directions
# show at different cutoffs (0.5, 0.75 etc) of BaseQTL. If we relax BaseQTL we hope to see the ones with higher beta turning green.
# don't need to show graphs, just want to see tables. purple/red stay and blues to become greens
# and make GUESS more stringent - 99. do we see stuff in the middle disappearing?
# by making GUESS stringent, do we have more concordance?
# look at the ~100 that are intersected across all 3

```

### Combine for CHi-C modality

```{r combine datasets: chic}
# remove the GWAS regions (not eQTLs)
base_chic_eQTLs <- base_chic[!int_id %like% "gwas"]

# Remove the SNPs in BaseQTL that intersect DpnII cut sites.
base_chic_eQTLs[, hg19Proxy_ID := tstrsplit(int_id, split = "_", keep = 1)]
to_remove <- dpn[all_dpn_ALT %like% "TRUE" | all_dpn_REF %like% "TRUE" | dpn_REF_in_baitRight == TRUE]
base_chic_dpn <- base_chic_eQTLs[!hg19Proxy_ID %in% to_remove$id]
base_chic_dpn[, `BaseQTL inverse norm. approx.` := 1 - norm_approx]

# Keep only the GUESS models that have only one SNP in the best model.
# split the betas and the SEs into the three modalities.
trip_distal[, c("chic_feat", "ATAC_peak", "Gene") := tstrsplit(`Triplet name`, split = "/")]
trip_one <- trip_distal[!`Post-proc. best model SNP` %like% ","]
trip_one[, c("GUESS_chic_beta", "GUESS_ATAC_beta", "GUESS_RNA_beta") := tstrsplit(`Post-proc. max mPPI post. E(Beta|Y)`, split = ",", type.convert = T)]
trip_one[, c("GUESS_chic_se", "GUESS_ATAC_se", "GUESS_RNA_se") := tstrsplit(`Post-proc. max mPPI post. SE(Beta|Y)`, split = ",", type.convert = T)]

# Generate Zscores in GUESS.
trip_one[, GUESS_Zscore_CHiC := GUESS_chic_beta/GUESS_chic_se]
trip_one[, GUESS_Zscore_ATAC := GUESS_ATAC_beta/GUESS_ATAC_se]
trip_one[, GUESS_Zscore_RNA := GUESS_RNA_beta/GUESS_RNA_se]


# Filter the GUESS data such that we only look at the interactions for encompassing DpnII fragment (to match BaseQTL analysis)
# need to make sure that the SNP is within the same fragment that we are testing.
# can use the dpn object.
dpn_small <- unique(dpn[, .(Chr, baitStart, baitEnd, baitID, id, hg38Proxy_ID)])
trip_one[, DpnID := tstrsplit(`Triplet name`, split = "_", keep = 1)]
trip_one[, DpnID := as.numeric(str_replace_all(DpnID, "X", ""))]
trip_one_dpn <- dpn_small[trip_one, on = c(id = "Post-proc. best model SNP", baitID = "DpnID"), nomatch = NULL]
trip_one_dpn[, hg38Proxy_pos := tstrsplit(hg38Proxy_ID, split = ":", keep = 2)]
# check that they're in the correct fragment
# trip_one_dpn[hg38Proxy_pos < baitEnd & hg38Proxy_pos > baitStart] # yes
# separately, to get an idea of how many "features" we test in BaseQTL, combine it with the dpn object.
base_chic_dpn_feat <- base_chic_dpn[dpn, on = c(hg19Proxy_ID = "id"), nomatch = NULL]
base_chic_dpn_feat[, ENSG := tstrsplit(int_id, split = "_", keep = 2)]
base_chic_dpn_feat[, chic_feat := paste(baitID, ENSG, Gene, sep = "_")]

# Calculate the Zscores in BaseQTL. NOTE- these are fold changes
base_chic_dpn[, Base_Zscore_CHiC := (2^log_mean_aFC)/(2^se_mean)]

# combine GUESS data with chic BaseQTL
both_chic <- trip_one_dpn[base_chic_dpn, on = c(id = "hg19Proxy_ID"), nomatch = NULL] # 199 lines

# reason for duplicate lines is that sometimes there is more than one ATAC peak tested for each CHiC interaction
# get columns for plotting
both_chic[, GUESS_FDR := `Post-proc. max mPPI` > 0.95]
both_chic[, BaseQTL_95 := `2.5%` < 0 & `97.5%` <0 | `2.5%` > 0 & `97.5%` >0 ]
both_chic[GUESS_FDR == T & BaseQTL_95 == T, significance := "95%_CI_in_both"]
both_chic[GUESS_FDR == T & BaseQTL_95 == F, significance := "95%_CI_in_GUESS"]
both_chic[GUESS_FDR == F & BaseQTL_95 == T, significance := "95%_CI_in_BaseQTL"]
both_chic[GUESS_FDR == F & BaseQTL_95 == F, significance := "neither_95%_CI"]
both_chic[AI_estimate < 0.4, AI_est := "<0.4"]
both_chic[AI_estimate > 0.4, AI_est := ">0.4"]
both_chic[is.na(AI_estimate), AI_est := "unavailable"]

# signify the post proc probability values
both_chic[, `GUESS Post-proc. best model post. prob. > 0.95` := `Post-proc. best model post. prob.` > 0.95]
fwrite(both_chic, file = "../comparisons/CHiC_BaseQTL_GUESS_Zscores.txt", sep = "\t", row.names = F, col.names = T, quote = F)
# get an idea of overall numbers
noSNPs <- length(unique(both_chic$hg38Proxy_ID))  
print(paste0("The number of SNPs that can be compared across GUESS and BaseQTL CHiC is ", noSNPs))
noModels <- length(unique(both_chic$`Triplet name`)) 
print(paste0("The number of models that will be compared across GUESS and BaseQTL CHiC is ", noModels))
# remember that there are more models than SNPs because we are testing more than one ATAC peak for some of them.

totalSNPs_BaseQTL <- length(unique(base_chic_dpn$hg19Proxy_ID))
totalSNPs_GUESS <- length(unique(trip_one$`Post-proc. best model SNP`))
totalCHiC_BaseQTL <- length(unique(base_chic_dpn_feat$chic_feat))
totalCHiC_GUESS <- length(unique(trip_one$chic_feat))
intersectCHiC <- length(unique(both_chic$chic_feat))

chic <- data.frame(Dataset = c("BaseQTL", "GUESS", "Intersect"), 
                  SNPs = c(totalSNPs_BaseQTL, totalSNPs_GUESS, noSNPs), 
                  CHiC_features = c(totalCHiC_BaseQTL, totalCHiC_GUESS, intersectCHiC))
print(chic)
# "Intersect" is the subset that we see in the following graphs.
```
### Chec how many of the BaseQTL SNPs we recalled with Triplet data
```{R}
both_chic_sig <- both_chic[Signif.99 == "yes"]


```


### plot the chic/GUESS modality
```{R CHiC plots}
# Plot the Zscores. Mark baseline allelic imbalance in case it is noticably important.
ggplot(both_chic, aes(x = `GUESS_Zscore_CHiC`, y = `Base_Zscore_CHiC`, col = significance, shape = AI_est)) +
  geom_point() +
  xlab("GUESS CHiC Zscore of beta") +
  ylab("BaseQTL CHiC Zscore of aFC") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle("GUESS versus BaseQTL CHi-C, Zscores")
  
ci_cor(both_chic$GUESS_Zscore_CHiC, both_chic$Base_Zscore_CHiC, method="spearman", type = "bootstrap", 
                                 R = 1200)

# Plot the effects. Mark baseline allelic imbalance in case it is noticably important.
ggplot(both_chic, aes(x = `GUESS_chic_beta`, y = `log_mean_aFC`, col = significance, shape = AI_est)) +
  geom_point() +
  xlab("GUESS CHiC beta") +
  ylab("BaseQTL CHiC log aFC") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle("GUESS versus BaseQTL CHi-C")
  
ci_cor(both_chic$GUESS_chic_beta, both_chic$log_mean_aFC, method="spearman", type = "bootstrap", 
                                 R = 1200)

```

Now, plot on different values of beta at mPPI >= 0.95 and define a cutoff.

```{R CHiC plots beta values}
quantile(both_chic$GUESS_chic_beta)
sig <- both_chic[GUESS_FDR == T]

mybetas <- seq(0, 0.1, 0.001)

toplot <- data.table()
for(beta in mybetas) {
  cutoff <- sig[abs(GUESS_chic_beta) >= beta]
  toTest <- unique(cutoff[, .(`Triplet name`, `Post-proc. FDR mPPI SNP`, GUESS_Zscore_CHiC, Base_Zscore_CHiC)])
  overlaps <- length(unique(toTest$`Post-proc. FDR mPPI SNP`))
  Triplets <- length(unique(toTest$`Triplet name`))
  res <- suppressWarnings(ci_cor(toTest$GUESS_Zscore_CHiC, toTest$Base_Zscore_CHiC, method="spearman", type = "bootstrap", 
                                 R = 1200))
  Rho <- res$estimate
  CI_2.5 <- res$interval[[1]]
  CI_97.5 <- res$interval[[2]]
  toplot_mybeta <- data.table(Beta_cutoff = beta, Spearman_Rho = Rho, CI_2.5 = CI_2.5, CI_97.5 = CI_97.5, 
                              No.SNPS = overlaps, No.Triplets = Triplets)
  toplot <- rbind(toplot, toplot_mybeta)
}

ggplot(toplot, aes(x = Beta_cutoff, y = Spearman_Rho, col = No.SNPS)) + 
  geom_line() + 
  ylab("Spearman Rho") +
  ggtitle("GUESS versus BaseQTL CHi-C Zscores, correlation at abs beta cutoffs with 95% CI") +
  scale_x_continuous(name ="Beta cutoff, CHi-C", breaks = seq(0, 0.1, 0.005)) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_color_gradientn(colors = colorspace::rainbow_hcl(50)) +
  geom_errorbar(aes(x = Beta_cutoff, y = Spearman_Rho, ymin = CI_2.5, ymax = CI_97.5))

### then plot at certain cutoffs
make_corPlot <- function(dt, cutoff) {
  toPlot <- dt[abs(GUESS_chic_beta) >= cutoff]
  ggplot(toPlot, aes(x = GUESS_Zscore_CHiC, y = Base_Zscore_CHiC, col = significance)) + 
  geom_point() + 
  xlab("GUESS CHiC Zscore of beta") +
  ylab("BaseQTL CHiC Zscore of aFC") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle(paste0("GUESS versus BaseQTL CHi-C, effect sizes at Beta >=", cutoff))
}

cutoff = 0
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.006
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.011
p <- make_corPlot(sig, cutoff)
print(p)




# use 0.006?

```


### Combine for ATAC modality

```{R combine datasets: ATAC}
# Get the normal approximation for the ATAC data
base_atac[, norm_approx := 1-pnorm(0, abs(log_mean_aFC), sd, lower.tail = FALSE)]
base_atac[, `BaseQTL inverse norm. approx.` := 1 - norm_approx]

# Again use only the GUESS models that have only one SNP in the best model.
# get the location of ATAC peaks to combine with BaseQTL
trip_one_peaks <- peak_locs[trip_one, on = c(Peak_name = "ATAC_peak")]
base_atac[, c("Chr_peakStart", "peakEnd", "hg19Proxy_ID") := tstrsplit(int_id, split = "-", type.convert = T)]
base_atac[, c("Chr", "peakStart") := tstrsplit(Chr_peakStart, split = ":", type.convert = T)]
base_atac[, chrom := paste0("chr", Chr)]

# Get the Zscores
base_atac[, Base_Zscore_ATAC := (2^log_mean_aFC)/(2^se_mean)]

# combine
both_atac <- base_atac[trip_one_peaks, on =c("chrom", "peakStart", "peakEnd", hg19Proxy_ID = "Post-proc. best model SNP"), nomatch = NULL]

# get columns for plotting
both_atac[, GUESS_FDR := `Post-proc. max mPPI` > 0.95]
both_atac[, BaseQTL_95 := `2.5%` < 0 & `97.5%` <0 | `2.5%` > 0 & `97.5%` >0 ]
both_atac[GUESS_FDR == T & BaseQTL_95 == T, significance := "95%_CI_in_both"]
both_atac[GUESS_FDR == T & BaseQTL_95 == F, significance := "95%_CI_in_GUESS"]
both_atac[GUESS_FDR == F & BaseQTL_95 == T, significance := "95%_CI_in_BaseQTL"]
both_atac[GUESS_FDR == F & BaseQTL_95 == F, significance := "neither_95%_CI"]
both_atac[AI_estimate < 0.4, AI_est := "<0.4"]
both_atac[AI_estimate > 0.4, AI_est := ">0.4"]
both_atac[is.na(AI_estimate), AI_est := "unavailable"]

# signify the post proc probability values
both_atac[, `GUESS Post-proc. best model post. prob. > 0.95` := `Post-proc. best model post. prob.` > 0.95]
fwrite(both_atac, file = "../comparisons/ATAC_BaseQTL_GUESS_Zscores.txt", sep = "\t", quote = F, row.names = F, col.names = T)
# get an idea of overall numbers
noSNPs <- length(unique(both_atac$hg19Proxy_ID))   
print(paste0("The number of SNPs that can be compared across GUESS and BaseQTL ATAC is ", noSNPs))
noModels <- length(unique(both_atac$`Triplet name`)) 
print(paste0("The number of models that will be compared across GUESS and BaseQTL ATAC is ", noModels))
# there are more models than SNPs because we are testing more than one CHiC peak for some of them.
totalSNPs_BaseQTL <- length(unique(base_atac$hg19Proxy_ID))
totalSNPs_GUESS <- length(unique(trip_one$`Post-proc. best model SNP`))
totalPeaks_BaseQTL <- length(unique(base_atac$Chr_peakStart))
totalPeaks_GUESS <- length(unique(trip_one$ATAC_peak))
intersectPeaks <- length(unique(both_atac$Peak_name))

atac <- data.frame(Dataset = c("BaseQTL", "GUESS", "Intersect"), 
                  SNPs = c(totalSNPs_BaseQTL, totalSNPs_GUESS, noSNPs), 
                  Peaks = c(totalPeaks_BaseQTL, totalPeaks_GUESS, intersectPeaks))
print(atac)

# "Intersect" is the subset that we see in the following graphs.
# note that there are many SNPs in BaseQTL ATAC because it was done genome-wide.

```

### plot the ATAC/GUESS modality
```{R ATAC plots}
# Plot the post proc posterior probability in GUESS versus normal approximation in BaseQTL. Mark baseline allelic imbalance in case it is noticably important.
ggplot(both_atac, aes(x = `GUESS_Zscore_ATAC`, y = `Base_Zscore_ATAC`, col = significance, shape = AI_est)) +
  geom_point() +
  xlab("GUESS Zscore ATAC beta") +
  ylab("BaseQTL Zscore ATAC aFC") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle("GUESS versus BaseQTL ATAC, Zscores")
  
ci_cor(both_atac$GUESS_Zscore_ATAC, both_atac$Base_Zscore_ATAC, method="spearman", type = "bootstrap", 
                                 R = 1200)

```

Define a cutoff for ATAC

```{R ATAC plots beta values}
quantile(both_atac$GUESS_ATAC_beta)
sig <- both_atac[GUESS_FDR == T]

mybetas <- seq(0, 0.1, 0.001)

toplot <- data.table()
for(beta in mybetas) {
  cutoff <- sig[abs(GUESS_ATAC_beta) >= beta]
  toTest <- unique(cutoff[, .(`Triplet name`, `Post-proc. FDR mPPI SNP`, GUESS_Zscore_ATAC, Base_Zscore_ATAC)])
  overlaps <- length(unique(toTest$`Post-proc. FDR mPPI SNP`))
  Triplets <- length(unique(toTest$`Triplet name`))
  res <- suppressWarnings(ci_cor(toTest$GUESS_Zscore_ATAC, toTest$Base_Zscore_ATAC, method="spearman", type = "bootstrap", 
                                 R = 1200))
  Rho <- res$estimate
  CI_2.5 <- res$interval[[1]]
  CI_97.5 <- res$interval[[2]]
  toplot_mybeta <- data.table(Beta_cutoff = beta, Spearman_Rho = Rho, CI_2.5 = CI_2.5, CI_97.5 = CI_97.5, 
                              No.SNPS = overlaps, No.Triplets = Triplets)
  toplot <- rbind(toplot, toplot_mybeta)
}

ggplot(toplot, aes(x = Beta_cutoff, y = Spearman_Rho, col = No.SNPS)) + 
  geom_line() + 
  ylab("Spearman Rho") +
  ggtitle("GUESS versus BaseQTL ATAC Zscores, correlation at abs beta cutoffs wiht 95% CI") +
  scale_x_continuous(name ="Beta cutoff, ATAC", breaks = seq(0, 0.1, 0.005)) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_color_gradientn(colors = colorspace::rainbow_hcl(50)) +
  geom_errorbar(aes(x = Beta_cutoff, y = Spearman_Rho, ymin = CI_2.5, ymax = CI_97.5))

### then plot at certain cutoffs
make_corPlot <- function(dt, cutoff) {
  toPlot <- dt[abs(GUESS_ATAC_beta) >= cutoff]
  ggplot(toPlot, aes(x = GUESS_Zscore_ATAC, y = Base_Zscore_ATAC, col = significance)) + 
  geom_point() + 
  xlab("GUESS ATAC Zscore of beta") +
  ylab("BaseQTL ATAC Zscore of aFC") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle(paste0("GUESS versus BaseQTL ATAC Zscore, effect sizes at Beta >=", cutoff))
}

cutoff = 0
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.001
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.01
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.02
p <- make_corPlot(sig, cutoff)
print(p)

### Use 0.001?
```

### Combine for RNA modality

```{R combine datasets: RNA}
# Get the normal approximation for the RNA data
base_rna[, norm_approx := 1-pnorm(0, abs(log2_aFC_mean), log2_aFC_sd, lower.tail = FALSE)]
base_rna[, `BaseQTL inverse norm. approx.` := 1 - norm_approx]

# Again use only the GUESS models that have only one SNP in the best model.
trip_one[, ENSG_gene := tstrsplit(`Triplet name`, split = "/", keep = 3)]
trip_one[, ENSG := tstrsplit(ENSG_gene, split = "_", keep = 1)]

# SNPs in the ATAC-seq data do not have chromosomes in the ID, so remove it in GUESS data as well.
trip_one[, SNP_noChr := sub(".*?:", "", `Post-proc. best model SNP`)]

# Get the Zscores
base_rna[, Base_Zscore_RNA := (2^log2_aFC_mean)/(2^log2_aFC_se_mean)]


# combine
both_rna <- base_rna[trip_one, on =c(Gene_id = "ENSG", SNP = "SNP_noChr"), nomatch = NULL]

# get columns for plotting
both_rna[, GUESS_FDR := `Post-proc. max mPPI` > 0.95]
both_rna[, BaseQTL_95 := `log2_aFC_2.5%` < 0 & `log2_aFC_97.5%` <0 | `log2_aFC_2.5%` > 0 & `log2_aFC_97.5%` >0 ]
both_rna[GUESS_FDR == T & BaseQTL_95 == T, significance := "95%_CI_in_both"]
both_rna[GUESS_FDR == T & BaseQTL_95 == F, significance := "95%_CI_in_GUESS"]
both_rna[GUESS_FDR == F & BaseQTL_95 == T, significance := "95%_CI_in_BaseQTL"]
both_rna[GUESS_FDR == F & BaseQTL_95 == F, significance := "neither_95%_CI"]
both_rna[min_AI < 0.4, AI_est := "<0.4"]
both_rna[min_AI > 0.4, AI_est := ">0.4"]
both_rna[is.na(min_AI), AI_est := "unavailable"]

# signify the post proc probability values
both_rna[, `GUESS Post-proc. best model post. prob. > 0.95` := `Post-proc. best model post. prob.` > 0.95]
both_rna2 <- copy(both_rna)
setnames(both_rna2, "log2_aFC_mean", "log_mean_aFC")
fwrite(both_rna2, file = "../comparisons/RNA_BaseQTL_GUESS_Zscores.txt", sep = "\t", quote = F, row.names = F, col.names = T)
# get an idea of overall numbers
noSNPs <- length(unique(both_rna$SNP))  
print(paste0("The number of SNPs that can be compared across GUESS and BaseQTL RNA is ", noSNPs))
noModels <- length(unique(both_rna$`Triplet name`)) 
print(paste0("The number of Triplet Features that will be compared across GUESS and BaseQTL RNA is ", noModels))
# there are more models than SNPs because we are testing more than one CHiC peak and ATAC peak for some of them.
totalSNPs_BaseQTL <- length(unique(base_rna$SNP))
totalSNPs_GUESS <- length(unique(trip_one$`Post-proc. best model SNP`))
totalGenes_BaseQTL <- length(unique(base_rna$Gene_id))
totalGenes_GUESS <- length(unique(trip_one$ENSG))
intersectGenes <- length(unique(both_rna$Gene_id))

rna <- data.frame(Dataset = c("BaseQTL", "GUESS", "Intersect"), 
                  SNPs = c(totalSNPs_BaseQTL, totalSNPs_GUESS, noSNPs), 
                  Genes = c(totalGenes_BaseQTL, totalGenes_GUESS, intersectGenes))
print(rna)

# "Intersect" is the subset that we see in the following graphs.
```

### plot the RNA/GUESS modality
```{R RNA plots}
# Plot the post proc posterior probability in GUESS versus normal approximation in BaseQTL. Mark baseline allelic imbalance in case it is noticably important.
ggplot(both_rna, aes(x = `GUESS_Zscore_RNA`, y = `Base_Zscore_RNA`, col = significance, shape = AI_est)) +
  geom_point() +
  xlab("GUESS Zscore RNA beta") +
  ylab("BaseQTL Zscore RNA aFC") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle("GUESS versus BaseQTL RNA, Zscores")
  
ci_cor(both_rna$GUESS_Zscore_RNA, both_rna$Base_Zscore_RNA, method="spearman", type = "bootstrap", 
                                 R = 1200)

```

```{R RNA plots beta values}
quantile(both_rna$GUESS_RNA_beta)
sig <- both_rna[GUESS_FDR == T]

mybetas <- seq(0, 0.1, 0.001)

toplot <- data.table()
for(beta in mybetas) {
  cutoff <- sig[abs(GUESS_RNA_beta) >= beta]
  toTest <- unique(cutoff[, .(`Triplet name`, `Post-proc. FDR mPPI SNP`, GUESS_Zscore_RNA, Base_Zscore_RNA)])
  overlaps <- length(unique(toTest$`Post-proc. FDR mPPI SNP`))
  Triplets <- length(unique(toTest$`Triplet name`))
  res <- suppressWarnings(ci_cor(toTest$GUESS_Zscore_RNA, toTest$Base_Zscore_RNA, method="spearman", type = "bootstrap", 
                                 R = 1200))
  Rho <- res$estimate
  CI_2.5 <- res$interval[[1]]
  CI_97.5 <- res$interval[[2]]
  toplot_mybeta <- data.table(Beta_cutoff = beta, Spearman_Rho = Rho, CI_2.5 = CI_2.5, CI_97.5 = CI_97.5, 
                              No.SNPS = overlaps, No.Triplets = Triplets)
  toplot <- rbind(toplot, toplot_mybeta)
}

ggplot(toplot, aes(x = Beta_cutoff, y = Spearman_Rho, col = No.SNPS)) + 
  geom_line() + 
  ylab("Spearman Rho") +
  ggtitle("GUESS versus BaseQTL RNA, correlation at abs beta cutoffs wiht 95% CI") +
  scale_x_continuous(name ="Beta cutoff, RNA", breaks = seq(0, 0.1, 0.005)) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_color_gradientn(colors = colorspace::rainbow_hcl(50)) +
  geom_errorbar(aes(x = Beta_cutoff, y = Spearman_Rho, ymin = CI_2.5, ymax = CI_97.5))

### then plot at certain cutoffs
make_corPlot <- function(dt, cutoff) {
  toPlot <- dt[abs(GUESS_RNA_beta) >= cutoff]
  ggplot(toPlot, aes(x = GUESS_Zscore_RNA, y = Base_Zscore_RNA, col = significance)) + 
  geom_point() + 
  xlab("GUESS RNA Zscore beta") +
  ylab("BaseQTL RNA Zscore aFC") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  ggtitle(paste0("GUESS versus BaseQTL RNA Zscore, effect sizes at Beta >=", cutoff))
}

cutoff = 0
p <- make_corPlot(sig, cutoff)
print(p)
cutoff = 0.004
p <- make_corPlot(sig, cutoff)
print(p)
cutoff = 0.01
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.015
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.02 # at this cutoff we already lose a significnat BaseQTL result
p <- make_corPlot(sig, cutoff)
print(p)

cutoff = 0.05 # at this cutoff we already lose a significnat BaseQTL result
p <- make_corPlot(sig, cutoff)
print(p)

### Use 0.004?
```

How many triplets and SNPs do we retain at these cutoffs?

```{R overall res}
# using:
chic_cutoff = 0.006
atac_cutoff = 0.001
expr_cutoff = 0.004

# in one snp model data
sig <- trip_one[`Post-proc. FDR mPPI` >= 0.95]
length(unique(sig$`Post-proc. FDR mPPI SNP`)) #  655
length(unique(sig$`Triplet name`)) # 1079
new <- sig[abs(GUESS_chic_beta) >= chic_cutoff & abs(GUESS_ATAC_beta) >= atac_cutoff & abs(GUESS_RNA_beta) >= expr_cutoff]
length(unique(new$`Post-proc. FDR mPPI SNP`)) #  484
length(unique(new$`Triplet name`)) # 713

# in all models, including multi SNPs
# have to put betas on new line for each sig SNP
trip_sig <- trip_distal[`Post-proc. FDR mPPI SNP` != ""]
# change every third delimiter to ";"
trip_sig[, "Post-proc. FDR mPPI post. mean Beta perSNP" := gsub("((?:[^,]+,){2}[^,]+),", "\\1;", `Post-proc. FDR mPPI post. mean Beta`) ]
# Do the same for the SNP, but every delimiter to ";"
trip_sig[, "Post-proc. FDR mPPI SNP single" := str_replace_all(`Post-proc. FDR mPPI SNP`, ",", ";")]

trip_sig[, ]

trip_newRows <- as.data.table(trip_sig %>% separate_rows(c(`Post-proc. FDR mPPI SNP single`, `Post-proc. FDR mPPI post. mean Beta perSNP`), sep = ";"))

trip_newRows[, c("GUESS_chic_beta", "GUESS_ATAC_beta", "GUESS_RNA_beta") := tstrsplit(`Post-proc. FDR mPPI post. mean Beta perSNP`, split = ",", type.convert = T)]

trip_newRows[, Gene := tstrsplit(`Triplet name`, split = "/", keep = 3)]

#trip_newRows[`Post-proc. best model size` > 1]

length(unique(trip_newRows$`Post-proc. FDR mPPI SNP single`)) #  828
length(unique(trip_newRows$`Triplet name`)) # 1329
length(unique(trip_newRows$Gene)) # 256

######### FINAL NUMBERS SHOWN HERE
sigNewNewrows <- trip_newRows[abs(GUESS_chic_beta) >= chic_cutoff & abs(GUESS_ATAC_beta) >= atac_cutoff & abs(GUESS_RNA_beta) >= expr_cutoff]
length(unique(sigNewNewrows$`Post-proc. FDR mPPI SNP`)) #  629
length(unique(sigNewNewrows$`Triplet name`)) # 931
length(unique(sigNewNewrows$Gene)) # 224
fwrite(sigNewNewrows, file = "../GUESS_DpnII_distal_betas.txt", sep = "\t", quote = F, row.names = F, col.names = T)
#########


######### Generate confusion matrix for each of the modalities at this cutoff 
make_confusion_matrix <- function(dt, col1, col2) {
  dt[get(col1) < 0, GUESS_sign := factor("negative")]
  dt[get(col1) > 0, GUESS_sign := factor("positive")]
  dt[get(col2) < 0, BASE_sign := factor("negative")]
  dt[get(col2) > 0, BASE_sign := factor("positive")]
  return(dt)
}

chic_both0.006 <- both_chic[abs(GUESS_chic_beta) >= 0.006 & GUESS_FDR == T]
chic_cm <- make_confusion_matrix(chic_both0.006, "GUESS_chic_beta", "log_mean_aFC")
confusionMatrix(data=chic_cm$GUESS_sign, reference = chic_cm$BASE_sign)
chisq.test(chic_cm$GUESS_sign, chic_cm$BASE_sign)

atac_both0.001 <- both_atac[abs(GUESS_ATAC_beta) >= 0.001 & GUESS_FDR == T]
atac_cm <- make_confusion_matrix(atac_both0.001, "GUESS_ATAC_beta", "log_mean_aFC")
confusionMatrix(data=atac_cm$GUESS_sign, reference = atac_cm$BASE_sign)
chisq.test(atac_cm$GUESS_sign, atac_cm$BASE_sign)

rna_both0.004 <- both_rna[abs(GUESS_ATAC_beta) >= 0.004 & GUESS_FDR == T]
rna_cm <- make_confusion_matrix(rna_both0.004, "GUESS_ATAC_beta", "log2_aFC_mean")
confusionMatrix(data=rna_cm$GUESS_sign, reference = rna_cm$BASE_sign)
chisq.test(rna_cm$GUESS_sign, rna_cm$BASE_sign)

```

```{R}
### How many of our regions were the effects going in the same direction?
library(pheatmap)
heat <- unique(sigNewNewrows[, .(`Triplet name`, `Post-proc. FDR mPPI SNP single`, GUESS_chic_beta, GUESS_ATAC_beta, GUESS_RNA_beta)])
heat[, ID := paste(`Triplet name`, `Post-proc. FDR mPPI SNP single`, sep = "_")]
heat[, c("Triplet name", "Post-proc. FDR mPPI SNP single") := NULL]
heat_df <- as.data.frame(heat)
rownames(heat_df) <- heat_df[,4]
heat_df2 <- heat_df[ -c(4) ]
heat_df2

library(RColorBrewer)
paletteLength <- 50
myColor <- colorRampPalette(c("#0072B2", "white", "orangered3"))(paletteLength)
myBreaks <- c(seq(min(heat_df2), 0, length.out = ceiling(paletteLength/2) + 1), 
                  seq(max(heat_df2)/paletteLength, max(heat_df2), length.out = floor(paletteLength/2)))
p <- pheatmap(heat_df2, cluster_rows = T, cluster_cols = F, color = myColor, breaks = myBreaks, show_rownames = F, clustering_distance_rows = "euclidean")
save_pheatmap_pdf <- function(x, filename, width=6, height=6) {
    stopifnot(!missing(x))
    stopifnot(!missing(filename))
    pdf(filename, width=width, height=height)
    grid::grid.newpage()
    grid::grid.draw(x$gtable)
    dev.off()
}
save_pheatmap_pdf(p, "../pheatmap_GUESS_betaFilt.pdf")

```

### Look at a confident example: exceeds the cutoffs in GUESS and is also signicant in BaseQTL
```{R find good example}
modelsKeep <- unique(sigNewNewrows[, .(`Triplet ID`, `Triplet name`, `Post-proc. FDR mPPI SNP`)])
modelsKeep


both_chic_sig <- both_chic[Signif.99 == "yes"]
Use <- both_chic_sig[modelsKeep, on = c("Triplet ID", "Triplet name", "Post-proc. FDR mPPI SNP"), nomatch = NULL]
Use

# interesting, the NFE2L3 locus is significant for CBX3 in GUESS! Can show this!
# I think CBX3 is a good one to show... or NAAA, both are good (Same beta directions)
#sigNewNewrows[`Triplet name` %like% "NFE2L3", .(GUESS_chic_beta, GUESS_ATAC_beta, GUESS_RNA_beta)]
sigNewNewrows[`Triplet name` %like% "OSCAR", .(`Post-proc. FDR mPPI SNP`, GUESS_chic_beta, GUESS_ATAC_beta, GUESS_RNA_beta)]
sigNewNewrows[`Triplet name` %like% "TFPT", .(GUESS_chic_beta, GUESS_ATAC_beta, GUESS_RNA_beta)]
# have to remember if we are looking at a dif feature, everuthing can be different

sigNewNewrows[`Triplet name` %like% "CBX3", .(`Triplet name`, `Post-proc. FDR mPPI SNP`, GUESS_chic_beta, GUESS_ATAC_beta, GUESS_RNA_beta)]
# CBX3: 5775056 baitID, 7:26095784:C:T hg38 proxy ID, 7:26135404:C:T hg19 proxy ID, 
# Peak_134539 is the one to lok at

```



```{R looking at distal SNPs}
### DISTAL SNPS NOW INCLUDED IN THE FIRST BIT OF CODE AT BEGINNING OF SCRIPT
# check how many of them were "truly distal" - for this need to go back to the triplet sets.
distal <- fread("~/Documents/analysis/BaseQTL_results/CHiC/round2_refBias/V02_final_filtered_eqtls_illumina_july19_hg38_Jan20.txt")
snpSets <- fread("~/Documents/analysis/from_leo/input/Helen_SNP_sets.txt")

# have to get in a table called something like distal_keep: set, gene
distal_small <- unique(distal[, .(hg19SNP_ID, hg19Proxy_ID, ENSG_ID, rep_Gene, Gene)])
intersect <- distal_small[snpSets, on =c(hg19Proxy_ID = "hg19_ID"), nomatch = NULL]
length(unique(intersect$set)) # 8971 (one was missing I remember)
length(unique(snpSets$set)) # 5741
intersect[, `:=` (ENSG_Gene = paste(ENSG_ID, Gene, sep = "_"), 
                  ENSG_repGene = paste(ENSG_ID, rep_Gene, sep = "_"))]
intersect_small <- unique(intersect[, .(set, ENSG_Gene, ENSG_repGene)])
fwrite(intersect, file = "../input/Helen_trulyDistalSets.txt", 
       sep = "\t", quote = F, col.names = T, row.names = F)

# now intersect the GUESS results
trip_newRows_distal1 <- trip_newRows[intersect_small, on = c(`Triplet ID` = "set", Gene = "ENSG_Gene"), nomatch = NULL]
trip_newRows_distal1[, ENSG_repGene := NULL]
trip_newRows_distal2 <- trip_newRows[intersect_small, on = c(`Triplet ID` = "set", Gene = "ENSG_repGene"), nomatch = NULL]
trip_newRows_distal2[, ENSG_Gene := NULL]

trip_newRows_distal <- unique(rbind(trip_newRows_distal1, trip_newRows_distal2))
length(unique(trip_newRows_distal$`Post-proc. FDR mPPI SNP single`)) #  828
length(unique(trip_newRows_distal$`Triplet name`)) # 1329
length(unique(trip_newRows_distal$Gene)) # 256

sigNewNewrows_distal <- trip_newRows_distal[abs(GUESS_chic_beta) >= chic_cutoff & abs(GUESS_ATAC_beta) >= atac_cutoff & abs(GUESS_RNA_beta) >= expr_cutoff]
length(unique(sigNewNewrows_distal$`Post-proc. FDR mPPI SNP`)) #  376
length(unique(sigNewNewrows_distal$`Triplet name`)) # 489
length(unique(sigNewNewrows_distal$Gene)) # 170
sigNewNewrows_distal[Gene %like% "THBS1"]
```